Direct observation of Feshbach enhanced \swave{} scattering of fermions

### 2015-09-12 draft1.tex
After Ian's post-torture comments. Did not yet figure out how to make subfigure labels be in the top left rather than bottom of figures.

\documentclass[12pt]{iopart}
\usepackage{iopams}
\usepackage{braket}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{cite}
\usepackage{algorithm}
\usepackage{algorithmic}
%\usepackage{ams}

\begin{document}
\newcommand{\swave}[0]{$\it{s}$-wave}
\newcommand{\pwave}[0]{$\it{p}$-wave}
\newcommand{\K}{$^{40}\rm{K}$}
\newcommand{\Rb}{$^{87}\rm{Rb}$}
\newcommand{\us}{$\rm{\mu s}$}
\newcommand{\mT}{$\rm{mT}$}
\newcommand{\ez}{$\bf{\hat{e}_z}$}
\newcommand{\ex}{$\bf{\hat{e}_x}$}
\newcommand{\um}{$\rm{\mu m}$}

@@

\section{Introduction}
Feshbach resonances are widely used for tuning the interaction strength in ultracold atomic gases. They have been particularly instrumental in the study of interactions and interaction-dependent processes in cold Fermi gases. In contrast to atomic Bose-Einstein condensates (BECs), where even weak interactions play a crucial role, for example giving rise to their characteristic Thomas-Fermi density profiles \cite{KetterleBEC}, the effects of interactions in Fermi gases are harder to observe. The density of Fermi clouds is typically $\sim1000$ times less than that of BECs\footnote{disprosium and erbium scitation goes here}, making it necessary to enhance the strength of interactions in order to observe significant interaction effects\cite{KetterleDFG}. The tunability of interactions provided by Feshbach resonances has allowed for creation of molecular Bose-Einstein condensates from Fermi gases \cite{Greiner03,Zwierlein03, Jochim03} as well as observation of the phase transition from the Bardeen-Cooper-Schrieffer (BCS) superconduting regime to the BEC regime at sufficiently low temperatures \cite{Bartenstein04, Bourdel04, Zwierlein04, Regal04}.
\par A Feshbach resonance occurs when the two-atom continuum energetically approaches a two-atom bound molecular state  \cite{Chin10, Timmermans99}. In experiment, the relative energy of the free atomic states in two hyperfine sublevels and the molecular state is defined by a bias magnetic field. Consequently, the Feshbach resonance can be accessed by changing the bias field. In the simple case where there are no inelastic two-body channels, such as for the \K{} resonance discussed in this work, the effect of the resonance on the scattering length between two free atoms is \cite{Chin10}
\begin{equation}
a(B)=a_{bg}\left(1-\frac{\Delta}{B-B_0}\right),
\label{feshbachEq}
\end{equation}
 $a_{bg}$ is the background scattering length, $\Delta$ is the width of the resonance, and $B_0$ is the field value at which the resonance occurs. The scattering length diverges at the resonance.
\par  The exact value of the resonant field $B_0$ is difficult to calculate analytically and is commonly computed via numerical models based on experimental input parameters \cite{Tiesinga93, Lysebo09, Gao11} or determined experimentally \cite{Inouye98, Cornish00}. Many experimental techniques have been used to characterize Feshbach resonances, including the observation of atom loss due to three-body inelastic scattering, measurement of re-thermalization timescales, and anisotropic expansion of the cloud upon release from a confining potential, all of which infer the elastic scattering cross section from collective behavior of the cloud \cite{Regal03,OHara02,Monroe93}.
\par Here we used direct scattering as a primary probe of the location and width of a Feshbach resonance. We collided pairs of ultra-cold Fermi gases and directly imaged the resulting \swave{} scattered atoms as a function magnetic field strength. This allowed us to observe the enhancement in scattering without relying on proxy effects. We measured the fraction of atoms scattered during the collision, and from this fraction deduced the resonant magnetic field  and width of the resonance.
%A similar technique has been used to characterize impurity scattering in BECs \cite{Chikkatur00}.
%\par The techniques developed in this experiment for observing Fermion scattering can be extended to engineering higher order partial wave interactions, as has been done for bosons \cite{Williams2012}.

In our dilute DFGs, even with the resonant enhancement of the scattering cross section, only a small fraction of the atoms scattered as the clouds passed through each other. This made direct detection of scattered atoms difficult due to detection uncertainty, disproportionately affecting regions of low atomic density. To optimize the signal-to-noise ratio (SNR) for low atom numbers, we absorption imaged with fairly long, high-intensity pulses --- a non-standard regime, where the atoms acquired a velocity during imaging and the resulting Doppler-shift was non-negligible.  Simulation of the absorption imaging process was necessary for an accurate interpretation of these images. With the simulation corrected images, we extracted the fraction of atoms scattered in our collision experiment.

This paper is in two parts. In the first, we study absorption imaging in the presence of a significant time-dependent Doppler shift and show how we use our results to interpret data. In the second, we describe our \swave{} scattering experiment and extract a measure of the location and width of the Feshbach resonance in \K{}.

\section{Absorption imaging in the presence of strong recoil induced detuning}
\label{sec:2}
Absorption imaging measures the shadow cast by an atomic ensemble in an illuminating probe laser beam with angular frequency $\omega_L$. This imaging technique relies on optical transitions between ground and excited atomic states. Such atomic transitions have an energy difference $\hbar\omega_0$, and a natural transition linewidth $\Gamma$. When interacting with a laser field an atom scatters photons from the laser field into the vacuum modes. In the two-level atom approximation, the rate of scattering is \cite{LCT}
\begin{equation}
\gamma_{sc}= \frac{\Gamma}{2} \frac{I/I_{\rm{sat}}}{1+(2\delta/\Gamma)^2 +I/I_{\rm{sat}}},
\label{eq:scatrate}
\end{equation}
where $\tilde{I}=I/I_{\rm{sat}}$ is the laser intensity in units of the saturation intensity, and $\tilde{\delta}=\delta/\Gamma$ is the detuning $\delta=\omega_L-\omega_0$ in units of the natural linewidth. 
\par An absorption image is obtained by shining an on- or near-resonant probe beam ($\delta\approx0$) onto the atomic cloud. Some of the light is scattered by the atoms, and the unscattered light, $I_f(x,y)$, is imaged onto a camera, as depicted in Fig. \ref{fig:absorptionIntor}(a) (top). The probe light is reapplied with the atoms absent to calibrate the intensity $I_0(x,y)$ of light unaffected by the atoms (bottom).
\begin{figure}
	\subfigure[]{\includegraphics*[scale=0.55]{Picture1a}}
	\subfigure[]{\includegraphics*{Picture1c.pdf}}
	\subfigure[]{\includegraphics*[scale=0.55]{Picture1b}}
\caption{Absorption imaging. (a) Near resonant probe light illuminates the atoms, and the transmitted light (containing a shadow of the atoms) is imaged on the camera. After the atoms depart an image of just the probe light is taken. (b)  The probe beam is partially absorbed as it traverses the cloud, and the intensity seen by atoms further along the imaging direction \ez{} is lowered.  (c) An atomic cloud illuminated by a probe light field absorbs photons from the probe and re-emits them in all directions. This process results in a net velocity of the cloud in the direction of the probe light as well as diffusive spreading in the transverse directions.  }
\label{fig:absorptionIntor}
\end{figure}

The intensity $I_f(x,y)$ is related to the number of atoms the light encountered. Consider the light as it travels along the imaging axis \ez{} through a 3-D atomic density profile $\rho(x,y,z)$. For the purposes of this discussion, we focus on a single pixel of the camera; and thus a single value of $I_0$ and $I_f$, and a single column of atoms, $\rho(z)$. We are not able to infer the entire atomic distribution, but for each pixel we may obtain the column density, $n = \int \rho\left(z\right) \,\mathrm{d}z$. As the light travels through a column of atoms, each atom scatters light according to Eq. (\ref{eq:scatrate}). Therefore, the atoms further along the imaging axis $z$ experience a reduced optical intensity due to attenuation of the laser field by the other atoms (Fig. \ref{fig:absorptionIntor}(b)). The intensity change from scattering as a function of $z$ is

@@

\frac{dI(z)}{dz}=-\hbar\omega_L\rho(z)\gamma_{sc}(z)=-\rho(z)\sigma_0\frac{I(z)}{1+D^2 +I/I_{\rm{sat}}},

@@

where $\sigma_0$ is the resonant scattering cross section and we define a scaled detuning $D=2\delta/\Gamma$. Integrating this equation yields a relation between the observed optical depth $OD=-\ln \left(I_f/I_0\right)$ and the atomic column density $n$ \cite{Reinaudi07} --- we call the column density deduced in this manner $\sigma_0 n^{\rm{(1)}}$:

@@

\sigma_0 n^{\rm{(1)}} = \left(1+D^2\right)OD+ \frac{I_0-I_f}{I_{\rm{sat}}}.

@@

When the probe intensity is much smaller than the saturation intensity, $I_0\ll I_{\rm{sat}}$, and the probe light is on resonance, $D=0$, the right hand side of Eq. (\ref{eq2}) reduces to the optical depth \cite{Reinaudi07}, giving the simple relationship $\sigma_0 n^{\rm{(0)}}=OD$. 
\par Equations (\ref{eq:dIdz})-(\ref{eq2}) neglect the atomic recoil momentum and its effect on the laser detuning \cite{Konstantinidis12}. When an atom absorbs a photon from the laser light field it acquires a momentum kick $\hbar k_R$ in the direction of the light field. The associated recoil velocity is $v_r=\hbar k_R/m$, where $\hbar k_R= h/\lambda$ is the recoil momentum of the light field with wavelength $\lambda$ and $m$ is the atomic mass. The re-emitted photon imparts a similar recoil momentum $\vec{p}_e$, but over many scattering events this momentum distribution averages to zero. Therefore the atom  will acquire an average velocity per photon $v_r$ along \ez{}. The variance of $\vec{p}_e$, however, is not zero, allowing the atoms to acquire some momentum transverse to the laser field. We will ignore this correction, but the effect of this on the atomic cloud is pictured in Fig. \ref{fig:absorptionIntor}(c).
\par This velocity parallel the light field Doppler shifts the atoms from resonance. After scattering $N$ photons an atom acquires an average velocity $N v_r$ and an additional detuning $\delta=k_R N v_r$. Therefore, even if the probe beam is initially on resonance with the atomic transition, we cannot neglect the detuning's  effect on the scattering rate. Furthermore, this detuning varies both with imaging time $t$ and with distance along the propagation direction $z$ (Fig. \ref{fig:expos}(a)). Thus, the intensity lost to the atoms also acquires a time dependence:

@@

\frac{dI(t,z)}{dz}=-\sigma_0 \rho(t,z) \frac{I(t,z)}{1+D(t,z)^2 +I(t,z)/I_{\rm{sat}}}, \label{eq3}

@@

where, if we assume that the atoms do not move significantly within the cloud during the imaging time, the scaled detuning $D$ can be expressed as

@@

D(t,z)=\frac{2 v_r}{\Gamma\hbar c \rho(t,z)}\int_0^t \frac{dI(z,\tau)}{dz}\,\mathrm{d}\tau; \label{eq4}

@@

\caption{(a) Dependence of velocity and detuning on position simulated for \K{} at three different imaging times and a probe intensity $I_0=0.8 I_{\rm{sat}}$. (b) Column densities deduced from optical depths obtained from recoil detuning corrected simulation of on-resonant imaging of $^{40}K$ atoms at probe intensity $I_0=0.8\, I_{\rm{sat}}$. The input column density $\sigma_0 n=1.6$. $\sigma_0 n^{\rm{(1)}}$ is the high probe intensity corrected optical density given by Eq. (\ref{eq2}). $\sigma_0 n^{\rm{(2)}}$ is the column density as expanded to second order in time, Eq. (\ref{eq:OD2}) \cite{LJLthesis}.}

@@

%\begin{align}
%\sigma_0 n^{\rm{(2)}}&=c_0+c_1t+c_2t^2 & {\text where} && c_0 &= \sigma_0 n^{\rm{(1)}}, 
%\end{align}
\sigma_0 n^{\rm{(2)}}=c_0+c_1t+c_2t^2
\sigma_0 n^{\rm{(2)}}= \sigma_0 n^{\rm{(1)}}+\frac{(kv_rt)^2}{3}\left[\frac{I_{\rm{sat}}}{I_f+I_{\rm{sat}}}-\frac{I_{\rm{sat}}}{I_0+I_{\rm{sat}}}+\mathrm{ln}\left(\frac{I_f+I_{\rm{sat}}}{I_0+I_{\rm{sat}}}\right)\right].
\label{eq:OD2}
As shown in  Fig. \ref{fig:expos}(b), the perturbative treatment is accurate to times up to the recoil time, defined as the time it would take a single atom to be detuned by half an atomic linewidth, but diverges after that point. To adequately correct for the recoil induced detuning of the atoms, we numerically simulated the imaging process to obtain $I_f$ as a function of imaging time, atomic density, and probe intensity.
%	\includegraphics*{figure2.pdf}
%\caption{Using time dependent $I_f$ values obtained from recoil detuning corrected simulation of on-resonant imaging of $^{40}K$ atoms at probe intensity $I_0=0.8\, I_{\rm{sat}}$, this graph shows the optical depths obtained by each model. The `true' optical depth is $\sigma_0 n=1.6$. $OD_1$ is the high probe intensity corrected optical depth given by Eq. (\ref{eq2}). $OD_2$ is the high probe intensity corrected and expanded to second order in time optical depth, Eq. (\ref{eq:OD2}) \cite{LJLthesis}. The recoil time is the time it takes for the cloud, on average, to become detuned by a linewidth $\Gamma$. Both models start to differ from the true value before a recoil time.  }
%\label{fig:ODcorrections}
\par In the following, we describe two versions of this simulation. First, we took a simplistic approach where the spatial distribution of atoms did not change appreciably during the imaging time: $vt\ll I_0/(\hbar \omega_L \gamma_{sc} \rho)$. We verified this approach in known limits and then cross-checked the validity of the static assumption. For realistic input parameters, we found this assumption to be invalid. We then developed a quasi-classical approach and allowed the atoms to move during the imaging time. While the atomic trajectories were wildly different than in the static atom approximation, we found that the predicted $OD$s only differed on the 0.5$\%$ level.

\subsection{Stationary atom model}
To solve Eqs. (\ref{eq3})-(\ref{eq4}), we simulated a 1-D Gaussian density distribution $\rho(z)$. We divided the cloud into spatial bins (the bin size was decreased until the result only differed at the $10^{-5}$ level from smaller bin sizes).  In this approximation, the number of atoms in each bin was time-independent.  The algorithm used is shown in Alg. (\ref{algorithm1}). We call the optical depth simulated by this algorithm the simulated optical depth $OD^{\rm{sim1}}$

\begin{algorithm}
\caption{Stationary atom model}
\label{algorithm1}
\begin{algorithmic}
\STATE $I[n=0,t]=I_0$ \COMMENT{$n$ is the bin index, $t$ is the time index, $I$ is in units of $I_{\rm{sat}}$}
\STATE $\delta[n, t=0]=0$ \COMMENT{light initially resonant, $\delta$ in units of $\Gamma/2$}
\STATE $H_f=0$ \COMMENT{Radiant fluence seen by camera after passing through cloud}
\FOR[loop over time steps]{$t=0$ to $t_f$}
 \FOR[loop over bins, N is total bin number]{$n=1$ to $N$}
 \STATE $A=\sigma_0\rho[n] dz$ \COMMENT{$dz$ is the size of spatial step}
 \STATE $B=v_r dt/(\hbar c \rho[n])$  \COMMENT{$dt$ is the size of the time step}
\STATE $I[n,t]=I[n-1,t] - A I[n-1,t]/(1+\delta[n,t-1]^2+I[n-1,t])$  \COMMENT{Eq. (\ref{eq3})}
\STATE $\delta[n,t]=\delta[n,t-1]+B\left(I[n-1,t]-I[n,t]\right)$  \COMMENT{Eq. (\ref{eq4})}
\ENDFOR
\STATE $H_f =H_f+ I[N,t]dt$ \COMMENT{collecting total fluence seen by the camera}
\ENDFOR
\STATE $OD^{\rm{sim1}}=-\ln{(H_f/I_0t_f)}$
\end{algorithmic}
\end{algorithm}

>>>>

\documentclass[12pt]{iopart}
\usepackage{iopams}
\usepackage{braket}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{cite}
\usepackage{algorithm}
\usepackage{algorithmic}
%\usepackage{caption}
%\usepackage{subcaption}
%\usepackage{harvard}
%\usepackage{amsmath}
%\usepackage{amssymb}

\begin{document}
\newcommand{\swave}[0]{$\it{s}$-wave}
\newcommand{\pwave}[0]{$\it{p}$-wave}
\newcommand{\K}{$^{40}\rm{K}$}
\newcommand{\Rb}{$^{87}\rm{Rb}$}
\newcommand{\us}{$\rm{\mu s}$}
\newcommand{\mT}{$\rm{mT}$}
\newcommand{\ez}{$\bi{e_z}$}
\newcommand{\ex}{$\bi{e_x}$}
\newcommand{\um}{$\rm{\mu m}$}
\renewcommand{\thefootnote}{\arabic{footnote}}

@@

\section{Introduction}
Feshbach resonances are widely used for tuning the interaction strength in ultracold atomic gases. They have been particularly instrumental in the study of interactions and interaction-dependent processes in cold Fermi gases. In contrast to atomic Bose-Einstein condensates (BECs), where even weak interactions play a crucial role, for example giving rise to their characteristic Thomas-Fermi density profiles \cite{KetterleBEC},  interractions must compete with the Fermi energy before becoming relevant. Practically speaking, the density of Fermi clouds is typically $\sim$1000 times less than that of BECs \footnote{This is not the case for recently realized erbium and dysprosium DFGs \cite{Aikawa14,Lu12}, where strong dipolar interactions are present.}, making it necessary to enhance the strength of interactions in order to observe significant interaction effects\cite{KetterleDFG}. The tunability of interactions provided by Feshbach resonances has allowed for creation of molecular Bose-Einstein condensates from Fermi gases \cite{Greiner03,Zwierlein03, Jochim03} as well as observation of the phase transition from the Bardeen-Cooper-Schrieffer (BCS) superconduting regime to the BEC regime at sufficiently low temperatures \cite{Bartenstein04, Bourdel04, Zwierlein04, Regal04}.
\par A Feshbach resonance occurs when a diatomic molecular state energetically approaches the two-atom continuum \cite{Chin10, Timmermans99}. In experiment, the relative energy of the free atomic states in two hyperfine sublevels and the molecular state is defined by a bias magnetic field. Consequently, the Feshbach resonance can be accessed by changing the bias field. In the simple case where there are no inelastic two-body channels, such as for the \K{} resonance discussed in this work, the effect of the resonance on the scattering length between two free atoms is \cite{Chin10}
\begin{equation}
a(B)=a_{\rm{bg}}\left(1-\frac{\Delta}{B-B_0}\right),
\label{feshbachEq}
\end{equation}
where $a_{\rm{bg}}$ is the background scattering length, $\Delta$ is the width of the resonance, and $B_0$ is the field value at which the resonance occurs. The scattering length diverges at the resonance.
\par  The exact value of the resonant field $B_0$ is difficult to calculate analytically and is commonly computed via numerical models based on experimental input parameters \cite{Tiesinga93, Lysebo09, Gao11} or determined experimentally \cite{Inouye98, Cornish00}. Many experimental techniques have been used to characterize Feshbach resonances, including the observation of atom loss due to three-body inelastic scattering, measurement of re-thermalization timescales, and anisotropic expansion of the cloud upon release from a confining potential, all of which infer the elastic scattering cross section from collective behavior of the cloud \cite{Regal03,OHara02,Monroe93}.
\par Here we used direct scattering as a primary probe of the location and width of a Feshbach resonance. We collided pairs of ultra-cold Fermi gases and directly imaged the resulting \swave{} scattered atoms as a function of magnetic field strength. This allowed us to observe the enhancement in scattering without relying on proxy effects. We measured the fraction of atoms scattered during the collision, and from this fraction deduced the resonant magnetic field  and the width of the resonance.
%A similar technique has been used to characterize impurity scattering in BECs \cite{Chikkatur00}.
%\par The techniques developed in this experiment for observing Fermion scattering can be extended to engineering higher order partial wave interactions, as has been done for bosons \cite{Williams2012}.

In our dilute DFGs, even with the resonant enhancement of the scattering cross section, only a small fraction of the atoms scattered as the clouds passed through each other. This made direct detection of scattered atoms difficult due to detection uncertainty that disproportionately affected regions of low atomic density. To optimize the signal-to-noise ratio (SNR) for low atom numbers, we absorption imaged with fairly long, high-intensity pulses --- a non-standard regime, where the atoms acquired a velocity during imaging and the resulting Doppler-shift was non-negligible.  Simulation of the absorption imaging process was necessary for an accurate interpretation of these images. Using the simulation-corrected images, we extracted the fraction of atoms scattered in our collision experiment.

This paper is divided into two parts. In the first, we study absorption imaging in the presence of a significant time-dependent Doppler shift and show how we use our results to interpret data. In the second, we describe our \swave{} scattering experiment and extract a measure of the location and width of the Feshbach resonance in \K{}.

\section{Absorption imaging in the presence of strong recoil induced detuning}
\label{sec:2}
Absorption imaging measures the shadow cast by an atomic ensemble in an illuminating probe laser beam with angular frequency $\omega_{\rm{L}}$. This imaging technique relies on optical transitions between ground and excited atomic states. Such atomic transitions have an energy difference $\hbar\omega_0$, and a natural transition linewidth $\Gamma$. When interacting with a laser field an atom scatters photons from the laser field into the vacuum modes. In the two-level atom approximation, the rate of scattering is \cite{LCT}
\begin{equation}
\gamma_{\rm{sc}}= \frac{\Gamma}{2} \frac{\tilde{I}}{1+4\tilde{\delta}^2 +\tilde{I}},
\label{eq:scatrate}
\end{equation}
where $\tilde{I}=I/I_{\rm{sat}}$ is the laser intensity in units of the saturation intensity, and $\tilde{\delta}=\delta/\Gamma$ is the detuning $\delta=\omega_{\rm{L}}-\omega_0$ in units of the natural linewidth. 
\par An absorption image is obtained by shining an on- or near-resonant probe beam ($\tilde{\delta}\ll1$) onto the atomic cloud. Some of the light is scattered by the atoms, and the unscattered light, $\tilde{I}_f(x,y)$, is imaged onto a camera, as depicted in Fig. \ref{fig:absorptionIntor}(a) (top). The probe light is reapplied with the atoms absent to calibrate the intensity $\tilde{I}_0(x,y)$ of light unaffected by the atoms (bottom).
\begin{figure}
	\subfigure[]{\includegraphics*[scale=0.5]{Picture1a}}
	\subfigure[]{\includegraphics*{Picture1c.pdf}}
	\subfigure[]{\includegraphics*[scale=0.55]{Picture1b}}
\caption{Absorption imaging. (a) Near resonant probe light illuminates the atoms, and the transmitted light (containing a shadow of the atoms) is imaged on the camera. A second image taken with no atoms provides a reference. (b)  The probe beam is partially absorbed as it traverses the cloud, and the intensity seen by atoms further along the imaging direction \ez{} is lowered.  (c) An atomic cloud illuminated by a probe light field absorbs photons from the probe and re-emits them in all directions. This process results in a net acceleration of the cloud in the direction of the probe light as well as diffusive spreading in the transverse directions.  }
\label{fig:absorptionIntor}
\end{figure}

The intensity $\tilde{I}_f(x,y)$ is related to the number of atoms the light encountered. Consider the light as it travels along the imaging axis \ez{} through a 3D atomic density profile $\rho(x,y,z)$. We focus on a single pixel of the camera: sensitive to a single column of atoms $\rho(z)$ giving  $\tilde{I}_0$ and $\tilde{I}_f$. Our task is to invert the problem and obtain a column density $n = \int \rho\left(z\right) \,\mathrm{d}z$ given $\tilde{I}_0$ and $\tilde{I}_f$. As the light travels through a column of atoms, each atom scatters light according to Eq. (\ref{eq:scatrate}). Therefore, the atoms further along the imaging axis $z$ experience a reduced optical intensity due to attenuation of the laser field by the other atoms (Fig. \ref{fig:absorptionIntor}(b)). The intensity change from scattering as a function of $z$ is

@@

\frac{d\tilde{I}(z)}{dz}=-\hbar\omega_{\rm{L}}\rho(z)\gamma_{sc}(z)=-\rho(z)\sigma_0\frac{\tilde{I}(z)}{1+4\tilde{\delta}^2 +\tilde{I}},

@@

where $\sigma_0$ is the resonant scattering cross section. Integrating this equation yields a relation between the observed optical depth $OD=-\ln \left(I_f/I_0\right)$ and the atomic column density $n$ \cite{Reinaudi07} --- we call the column density deduced in this manner $\sigma_0 n^{\rm{(1)}}$:

@@

\sigma_0 n^{\rm{(1)}} = \left(1+4\tilde{\delta}^2\right)OD+ \tilde{I}_0-\tilde{I}_f.

@@

When the probe intensity is much smaller than the saturation intensity, $\tilde{I}_0\ll1$, and the probe light is on resonance, $\tilde{\delta}=0$, the right hand side of Eq. (\ref{eq2}) reduces to the optical depth \cite{Reinaudi07}, giving the simple relationship $\sigma_0 n^{\rm{(0)}}=OD$. 
\par Equations (\ref{eq:dIdz})-(\ref{eq2}) neglect the atomic recoil momentum and its effect on the laser detuning \cite{Konstantinidis12}. When an atom absorbs a photon from the laser light field it acquires a momentum kick $\hbar  k_{\rm{r}}$ in the direction of the light field. The associated recoil velocity is $v_{\rm{r}}=\hbar k_{\rm{r}}/m$, where $m$ is the atomic mass and $\hbar k_{\rm{r}}= h/\lambda$ is the recoil momentum from the laser with wavelength $\lambda$. Each re-emitted photon imparts a similar recoil momentum $\bi{p}_e$, but over many scattering events this momentum distribution averages to zero. Therefore the atom  will acquire an average velocity per photon $v_{\rm{r}}$ along \ez{}. The variance of $\bi{p}_e$, however, is not zero, allowing the atoms to acquire some momentum transverse to the laser field. While we ignore this correction, it results in the reduction of spatial resolution in the final image and its effect on the atomic cloud is pictured in Fig. \ref{fig:absorptionIntor}(c).
\par The average atom velocity parallel the light field after scattering $N$ photons is $N v_{\rm{r}}$ and it is Doppler shifted  $\delta= k_{\rm{r}} N v_{\rm{r}}$ from resonance. For any probe intensity, there is an imaging time that we call the recoil time $t_{\rm{r}}$ after which  $\delta>\Gamma/2$ and, even if the probe beam is initially on resonance with the atomic transition, we cannot neglect the detuning's  effect on the scattering rate. Furthermore, this detuning varies both with imaging time $t$ and with distance along the propagation direction $z$ (Fig. \ref{fig:expos}(a)). Thus, the laser's spatially varying intensity profile in the atomic cloud also depends on time:

@@

\frac{d\tilde{I}(t,z)}{dz}=-\sigma_0 \rho(t,z) \frac{\tilde{I}(t,z)}{1+4\tilde{\delta}(t,z)^2 +\tilde{I}(t,z)}. \label{eq3}

@@

Assuming that the atoms do not move significantly during the imaging time (we will remove this assumption shortly), the dimensionless detuning is

@@

\tilde{\delta}(t,z)=\frac{ k_{\rm{r}} v_{\rm{r}}}{2\sigma_0 \rho(t,z)}\int_0^t \frac{d\tilde{I}(z,\tau)}{dz}\,\mathrm{d}\tau; \label{eq4}

@@

\caption{(a) Dependence of velocity and detuning on position simulated for \K{} at three different imaging times and a probe intensity $\tilde{I}_0=0.8$. (b) Column densities deduced from optical depths obtained from recoil detuning corrected simulation of on-resonant imaging of $^{40}K$ atoms at probe intensity $\tilde{I}_0=0.8$. The input column density $\sigma_0 n=1.6$. $\sigma_0 n^{\rm{(1)}}$ is the high probe intensity corrected column density given by Eq. (\ref{eq2}). $\sigma_0 n^{\rm{(2)}}$ is the column density as expanded to second order in time, Eq. (\ref{eq:OD2}).}

@@

\begin{eqnarray}
&\sigma_0 n^{\rm{(2)}}  =  c_0+c_1t+c_2t^2, \mbox{ where } \\
 &c_0= \sigma_0 n^{\rm{(1)}}, c_1=0, c_2=\frac{( k_{\rm{r}} v_{\rm{r}})^2}{3}\left[\frac{1}{\tilde{I}_f+1}-\frac{1}{\tilde{I}_0+1}+\mathrm{ln}\left(\frac{\tilde{I}_f+1}{\tilde{I}_0+1}\right)\right]
\label{eq:OD2}
\end{eqnarray}
As shown in  Fig. \ref{fig:expos}(b), the perturbative treatment is accurate to times up to the recoil time after which it begins to diverge. To adequately correct for the recoil induced detuning of the atoms, we numerically simulated the imaging process to obtain $\tilde{I}_f$ as a function of imaging time, atomic density, and probe intensity.
%\begin{figure}
%	\includegraphics*{figure2.pdf}
%\caption{Using time dependent $I_f$ values obtained from recoil detuning corrected simulation of on-resonant imaging of $^{40}K$ atoms at probe intensity $I_0=0.8\, I_{\rm{sat}}$, this graph shows the optical depths obtained by each model. The `true' optical depth is $\sigma_0 n=1.6$. $OD_1$ is the high probe intensity corrected optical depth given by Eq. (\ref{eq2}). $OD_2$ is the high probe intensity corrected and expanded to second order in time optical depth, Eq. (\ref{eq:OD2}) \cite{LJLthesis}. The recoil time is the time it takes for the cloud, on average, to become detuned by a linewidth $\Gamma$. Both models start to differ from the true value before a recoil time.  }
%\label{fig:ODcorrections}
%\end{figure}
\par In the following, we describe two versions of this simulation. First, we took a simplistic approach where the spatial distribution of atoms did not change appreciably during the imaging time: $vt\ll I_0/(\hbar \omega_{\rm{L}} \gamma_{\rm{sc}} \rho)$. We verified this approach in known limits and then cross-checked the validity of the static assumption. For realistic input parameters, we found this assumption to be invalid. We then developed a quasi-classical approach and allowed the atoms to move during the imaging time. While the atomic trajectories were wildly different than in the static atom approximation, we found that the predicted $OD$s only differed on the 0.5$\%$ level.

\subsection{Stationary atom model}
To solve Eqs. (\ref{eq3})-(\ref{eq4}), we divided the cloud into spatial bins.  In this approximation, the number of atoms in each bin was time-independent.  The algorithm used is shown in Alg. (\ref{algorithm1}), in which we took a Gaussian profile for our initial density distribution. We call the optical depth simulated by this algorithm the simulated optical depth $OD^{\rm{sim1}}$.

\begin{algorithm}
\caption{Stationary atom model}
\label{algorithm1}
\begin{algorithmic}
\STATE $I[n=0,t]=I_0$ \COMMENT{$n$ is the bin index, $t$ is the time index, $I$ is in units of $I_{\rm{sat}}$}
\STATE $\delta[n, t=0]=0$ \COMMENT{light initially resonant, $\delta$ in units of $\Gamma/2$}
\STATE $H_f=0$ \COMMENT{Radiant fluence seen by camera after passing through cloud}
\FOR[loop over time steps]{$t=0$ to $t_f$}
 \FOR[loop over bins, N is total bin number]{$n=1$ to $N$}
 \STATE $A=\sigma_0\rho[n] dz$ \COMMENT{$dz$ is the size of spatial step}
 \STATE $B=v_{\rm{r}} dt/(\hbar c \rho[n])$  \COMMENT{$dt$ is the size of the time step}
\STATE $I[n,t]=I[n-1,t] - A I[n-1,t]/(1+\delta[n,t-1]^2+I[n-1,t])$  \COMMENT{Eq. (\ref{eq3})}
\STATE $\delta[n,t]=\delta[n,t-1]+B\left(I[n-1,t]-I[n,t]\right)$  \COMMENT{Eq. (\ref{eq4})}
\ENDFOR
\STATE $H_f =H_f+ I[N,t]dt$ \COMMENT{collecting total fluence seen by the camera}
\ENDFOR
\STATE $OD^{\rm{sim1}}=-\ln{(H_f/I_0t_f)}$
\end{algorithmic}
\end{algorithm}

\par We checked the validity of our simulation in the limits where the problem is analytically solvable. In the limit where the probe intensity is much weaker than the saturation intensity, $\tilde{I}_0\ll 1$, the atoms' velocities are hardly changed, and Eq.(\ref{eq3}) reduces to
\begin{eqnarray}
\frac{d\tilde{I}(z)}{dz}&=-\rho\sigma_0\tilde{I}(z), \mbox{ from which we recover the analytic form }\\
\sigma_0 n^{\rm{(0)}} &= -\ln\tilde{I}_0/\tilde{I}_f. \label{eq6}
\end{eqnarray}
In the limit where the probe intensity is much larger than the saturation intensity, $\tilde{I}_0\gg \tilde{\delta}$, even far detuned atoms will scatter light at their maximum rate. The time dependence of the detuning can thus be neglected, and Eq. (\ref{eq3}) becomes
\begin{eqnarray}
\frac{d\tilde{I}(z)}{dz}&=-\rho\sigma_0, \mbox{ which integrates to }\\
\sigma_0 n &= \tilde{I}_0 - \tilde{I}_f. \label{eq8}
\end{eqnarray}
We recognize the right hand sides of Eq. (\ref{eq6}) and Eq. (\ref{eq8}) as the two terms in Eq. (\ref{eq2}). Thus, as shown in  Fig. \ref{fig:IsatLimits}, $OD^{\rm{sim1}}$  coincides with the optical depth as predicted by Eq. (\ref{eq2}) in both the small and large probe intensity limits.
\begin{figure}
	\includegraphics{figure3.pdf}
\caption{Optical depth as a function of probe intensity as predicted by the simulation (blue symbols) and by Eq. (\ref{eq2}) (green curves), for three different imaging times. As expected, the predictions agree in both the high and low intensity limits, and differ for probe intensities comparable to the saturation intensity and longer imaging times. }
\label{fig:IsatLimits}
\end{figure}
\par We used the results of this simulation to check the self-consistency of the  stationary atom assumption, i.e. the distance traveled by the atoms (as deduced from integrating the acquired recoil velocity over the imaging time) is less than the bin size. As can be seen from Fig. \ref{fig:simTests}(a), not only do the atoms travel more than the bin size, but they travel far beyond the initial extent of the cloud. Moreover, owing to the higher initial scatter rate, the back of the cloud overtakes the front for long imaging times. Thus, the atomic distribution as a function of position changes dramatically during the imaging pulse, and the stationary assumption is invalid.
\begin{figure}
	\subfigure[]{\includegraphics{figure4.pdf}}
	\subfigure[]{\includegraphics{figure5.pdf}}
\caption{(a) Position of atoms as a function of imaging time for atoms in the first (solid green), middle (dashed red), and last (dotted blue) bins of the simulated density distribution for an initial cloud 50 \um{} in extent. The probe intensity used in this calculation was $1.2\, I_{\rm{sat}}$, and the column density was $\sigma_0 n=1.6$. (b) The velocity of a single composite atom as a function of probe intensity for various imaging times. Simulation data (dots) and numerical  solutions of Eq. (\ref{eq10}) (lines) are in agreement.}
\label{fig:simTests}
\end{figure}
